Effects of the body force on the pedestrian and evacuation dynamics

### 2020-02-05 paper.tex
correcciones en la primer parte de corridor

We may summarize this Section as follows. We explored the parameter space along $v_d$ and $k_n$ for the bottleneck situation. Varying $v_d$ (at a fixed $k_n$) explores the collective behavior for different $(\mathcal{A},\mathcal{K}_c)$ pairs. We realized that increasing $\mathcal{K}_c$ only improves the evacuation performance through the bottleneck.

@@

In this section, we present the results corresponding to the corridor geometry. We show the effects of modifying the body force coefficient $k_n$ in the dynamics.\\

In the same way as with the bottleneck geometry, we computed the contact network to study the topological differences that arise from increasing the $k_n$ value. Fig.~\ref{degree_dens} shows the mean degree as a function of the global density for different $k_n$ values. When the density is low,  the degree is zero because the pedestrians do not touch each other. When the density is around 4.5, a few pedestrians start to touch each other increasing the mean degree. As the density increases, the mean degree increases until reaching an asymptotic behavior as approaching degree six. Degree six corresponds to the densest packing possible for identical hard circles; the result is a hexagonal packing arrangement.\\

It is important to notice that in Fig. \ref{degree_dens}, for a given value of density, the stiffer the pedestrians, the higher the degree. In the bottleneck geometry, the opposite thing happens. The stiffer the pedestrians, the lower the mean degree.
\\

The mean overlap increases as the global density increases (see Fig. \ref{overlap_dens}). For a given density, there is an inverse relation between stiffness and mean overlap. This result is similar to the overlap vs. desired velocity obtained for bottleneck geometry. The higher the value of $k_n$, the higher the repulsion of the pedestrian-pedestrian interaction since the body force is acting on the direction $n_{ij}$.\\

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_dens_multi_kn.eps}\label{degree_dens}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_dens_multi_kn.eps}\label{overlap_dens}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the global density for different $k_n$ values. (b) mean overlap as a function of the global density. The mean values are averages over all the pedestrians and over time once the system reaches the stationary state. Both measurements correspond to corridor geometry with desired velocity $v_d=1$. Color online only.}
\label{network_corridor}
\end{figure}

As was mentioned above, increasing the $k_n$ value of the body force increases the stiffness of pedestrians. Increasing the stiffness reduces the overlap (both in the bottleneck and corridor). Regarding the network connectivity, increasing the stiffness reduces the mean degree in a bottleneck but increases the mean degree in a corridor. Fig.~\ref{network_corridor} shows the contact network corresponding to the corridor at a particular timestep ($t=50$~s) and at a particular global density ($\rho=7$). Fig.~\ref{network_d6_kn0} corresponds to $k_n=0$ while Fig.~\ref{network_d6_knE5} corresponds to $k_n=1.2$~E5. This is an example to show that increasing the stiffness increases the number of contacts between pedestrians (the network connectivity). Notice that there are many more pedestrians with degree 6 in the case of $k_n=1.2$~E5 than the case of $k_n=0$. \\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor. The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2$~E5. Color online only.}
\label{network_corridor}
\end{figure}

>>>>

We may summarize this Section as follows. We explored the parameter space along $v_d$ and $k_n$ for the bottleneck situation. This is similar to explore the unitless space $(\mathcal{A},\mathcal{K}_c)$. Soft pedestrians attain a FIS or FIS+FIF evacuation pattern but, stiff pedestrians (for any fixed $\mathcal{A}$ value within the explored range) exhibit a single FIF pattern. Stiffness affects either the presence of the blocking clusters and the pedestrians overlap. The less they compress, the less intense becomes the sliding friction among them. Additionally, the fewer the blocking clusters the easier they get out. \\

@@

We present in this section the results corresponding to the corridor geometry. We show the effects of modifying the body force coefficient $k_n$ on the collective dynamics.\\

We computed the contact network in the same way in the bottleneck geometry. Fig.~\ref{degree_dens} shows the mean degree as a function of the global density for different $k_n$ values (see caption for details). The degree is zero because the pedestrians do not touch each other. When the density reaches 4.5, a few pedestrians start to touch each other increasing the mean degree. As the density continuous increasing, the mean degree increases towards the asymptotic degree six. Degree six corresponds to the maximum packing density for identical hard circles.\\

Fig. \ref{degree_dens} exhibits a somewhat opposite behavior with respect to the bottleneck geometry, for a given value of density. Here, the stiffer the pedestrians the higher the degree. In the bottleneck geometry, however, the stiffer the pedestrians, the lower the mean degree. Besides, Fig. \ref{overlap_dens} shows that the mean overlap increases as the global density increases. This behavior appears similar to the bottleneck situation, in the sense that the higher the value of $k_n$ the less they overlap. The same reasoning in Section \ref{bottleneck} apply to this situation.  \\

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_dens_multi_kn.eps}\label{degree_dens}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_dens_multi_kn.eps}\label{overlap_dens}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the global density for different $k_n$ values. (b) mean overlap as a function of the global density. The global density is the total number of pedestrians per unit area. The mean values are averages over all the pedestrians and over time once the system reached the stationary state. Both measurements correspond to corridor geometry with desired velocity $v_d=1$. Color online only.}
\label{network_corridor}
\end{figure}

In order to get a better insight on how the pedestrians contact to each other, we present in Fig.~\ref{network_corridor} two snapshots of the corridor at the stationary situation. Fig.~\ref{network_d6_knE5} corresponds to $k_n=1.2$~E5. It can be seen that increasing ther stiffness increases the number of contacts between pedestrians (say, the network connectivity). The number of nodes attaining degree 6 is far larger in the case of $k_n=1.2\,$E5 than the case of $k_n=0$. \\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2$~E5. The friction coefficient and the other SFM parameters are the same as in Section \ref{bottleneck}. Color online only.}
\label{network_corridor}
\end{figure}
